Easy as Pi: A Network Coding Raspberry Pi Testbed

### 2016-06-05 schemes.tex
proofreading. Added info on Budget and density for TSNC

In this section, we present a description of the considered coding schemes
measured with our the \ac{Raspi} testbed. We introduce a definition
for the primitive coding operations, e.g. encoding, decoding and recoding
(where applies) for each coding scheme. Later, we address particular schemes
which are obtained by modifying the basic coding operations that provide
better processing speeds, particularly relevant for the \ac{Raspi}.
Finally, we include a review of the

@@

\ac{RLNC} is an example of intra-session \ac{NC}. Here, data symbols from a
single flow are combined with each other. In this type of network coding,
all the original data packets $P_j, j \in [1,g]$, each of $B$ bytes, are used
to create coded packets. In the following subsections, we describe the
basic functionalities of \ac{RLNC}.

\subsubsection{Encoding}
In \ac{RLNC}, any coded packet is a linear combination from all
the original packets. For the coding scheme, packets are seen as
algebraic entities formed as a sequence of elements from $GF(q)$,
which is a \ac{GF} of size $q$. Later, each original packet is
multiplied by a coding coefficient from $GF(q)$. The coding coefficients
are picked uniformly at random from the \ac{GF} by the encoder. To
perform the multiplication of a packet by a coding coefficient, the
coefficient is multiplied for each of the elements in the
concatenation that composes an original packet, preserving the
concatenation. Later, all resulting packets are added within the
\ac{GF} arithmetics together to generate a coded packet. Thus, a
coded packet can be written as:
%
\begin{align} \label{eq:coded_packet}
C_i  &= \bigoplus_{j=1}^{g} v_{ij} \otimes P_j ,\ \forall i \in [1,\ldots)
\end{align}

@@

For indicating a receiver which packets were utilized
to create a coded one, a simple yet versatile choice is to append
its coding coefficients as a header of the coded packet. Hence, an
amount of overhead is included in every coded packet given that we
need to provide some signalling needed for decoding. The coding
coefficients overhead amount for packet $i$, $|v_i|$, can be quantified
as:
%
\begin{align} \label{eq:coded_packet_coef}
|v_i| &= \sum_{j=1}^{g} |v_{ij}| = g \times \lceil \log_{2} (q)
\rceil ~ [\mathrm{bits}]
\end{align}

\subsubsection{Decoding}
\label{sssec:decoding}
To be able to decode a batch of $g$ packets, a \ac{l.i.} set of $g$
coded packets, $C_i,\ i \in [1,g]$ is required at a decoder.
Once this set has been collected for a decoder,
the original packets can be found by computing the solution of a system
of linear equations using \ac{GF} aritmethics. Thus, we define
$\textbf{C} = \left[C_1 \ldots C_g \right]^T$,
$\textbf{P} = \left[P_1 \ldots P_g \right]^T$ and the coding matrix
$\textbf{V}$ that collects the coding coefficients for each of the $g$
coded packets, as follows:
%
\begin{align} \label{eq:coding_matrix}
\textbf{V} =
\left[
\begin{array}{c}
        v_1    \\ \hline
        \vdots \\ \hline
        v_g    \\
\end{array}
\right]
=
\left[
\begin{array}{ccc}
        v_{11} & \ldots & v_{1g} \\
        \vdots  & \ddots & \vdots  \\
        v_{g1} & \ldots & v_{gg} \\
\end{array}
\right]
\end{align}
%
Algebraically, decoding simplifies to finding the inverse of
$\textbf{V}$ in the linear system
$\textbf{C} = \textbf{V} \textbf{P}$, which can achieved using
efficient Gaussian elimination techniques \cite{fragouli2006network}.
On real applications, decoding is performed on-the-fly, e.g. the pivots are
computed as packets are progressively received, in order to minimize the
computation delay for each step.

@@

In \eqref{eq:recoded_packet}, $w_{ij}$ is the coding coefficient used
in the $i$-th recoded packet and assigned to multiply a previously coded
packet $C_j$. These coding coefficients are again uniformly and randomly chosen
from $GF(q)$. Still, each of the $w_{ij}$ are not the appended recoding
coefficients for signalling since the they will only retrieve previously
coded packets. The analytical expression for the appended coefficients
after recoding is obtained by calculating recoded packets as function of the
original packets. Therefore, we define the local coding matrix $\textbf{W}$
in the same way as it was made for $\textbf{V}$. Thus, the local coding matrix
can be written as:
%
\begin{align} \label{eq:local_coding_matrix}
\textbf{W} =
\left[
\begin{array}{c}
        w_1    \\ \hline
        \vdots \\ \hline
        w_g    \\
\end{array}
\right]
=
\left[
\begin{array}{ccc}
        w_{11} & \ldots & w_{1g} \\
        \vdots  & \ddots & \vdots  \\
        w_{g1} & \ldots & w_{gg} \\
\end{array}
\right]
\end{align}
%
With the definitions from \eqref{eq:coding_matrix} and
\eqref{eq:local_coding_matrix}, the recoded packets
$\textbf{R} = \left[R_1 \ldots R_g \right]^T$ are written as
$\textbf{R} = (\textbf{W} \textbf{V}) \textbf{P}$. Here, we recognize
the relationship between original and recoded packets. The resulting
coding matrix is simply the multiplication of matrices $\textbf{W}$ and
$\textbf{V}$. Denoting ${(\textbf{W} \textbf{V})}_{ij}$ as the element
in the $i$-th row and $j$-th column of $(\textbf{W} \textbf{V})$, this
term is the resulting coding coefficient used to create $R_i$ after
encoding the original packet $P_j$ recoding it locally in an intermediate
node. Finally, the appended coefficients for $R_i$ are
${(\textbf{W} \textbf{V})}_{ik}$ with $k \in [1,g]$. By doing some
algebra, each ${(\textbf{W} \textbf{V})}_{ij}$ term can be verified to be
computed as:
%
\begin{align}
\label{eq:appended_coded_coefficients}
{(\textbf{W} \textbf{V})}_{ij} = \sum_{k=1}^{g} w_{ik} v_{kj}\ ,\ \forall i,j \in [1,g] \times [1,g]
\end{align}
%
This procedure is carried by all the recoders in a given network, therefore
allowing any decoder to compute the original data after Gaussian elimination,
regardless of the amount of times where recoding was performed. Similarly
as with the encoding operation, any decoder that collects a set of
$g$ \ac{l.i.} recoded packets with their respective coefficients, will
be able to decode the data as mentioned before in Section \ref{sssec:decoding}.

@@

In \ac{SRLNC}, instead of considering all the packets to create a coded
packet as in \ac{RLNC}, for this coding scheme an encoder sets more
coding coefficients to zero when generating a coded packet with the purpose
of reducing the overall processing. Decoding is the same as in \ac{RLNC}, but
given that the coding matrices are now sparse, we will have less operations
performed in the decoding process. Recoding, although theoreticaly possible,
is omitted in practical systems since it is too complex to keep packets sparse
after recoding [\textbf{reference}]. In what follows, we
describe the coding scheme with a basic method to produce sparse coded
packets and later we indicate a second one, giving reference to their
implementation performance.

@@

To achieve a desired average coding density in the coefficients, for each
of them, we utilize a set of Bernoulli random variables all with parameter
$d$ as it success probability, i.e. $\mathbb{B}_j \sim Bernoulli(d)\ ,\ \forall
j \in [1,g]$. In this way, we can represent a coded packet in \ac{SRLNC} as:
%
\begin{align} \label{eq:sparse_coded_packet}
    C_i  &= \bigoplus_{j=1}^{g} \mathbb{B}_j v_{ij} \otimes P_j ,\ v_{ij}
    \neq 0\
    \forall i \in [1,\ldots) ,\ \
    d \in
    \begin{cases}
        {(0,0.5]} & ,\ q = 2 \\
        {(0,1]}   & ,\ q > 2
    \end{cases}
\end{align}
%
In \eqref{eq:sparse_coded_packet}, we have the requirement for the coding
coefficient to not be zero, since we want to ensure a coding coefficient
is generate for any random trial where $\mathbb{B}_j = 1$. Therefore, in our
implementation of \ac{SRLNC} we exclude the zero element and then pick
uniformly distributed random elements from $GF(q)-\{ 0 \}$. Also, we have
specified a dependency on the field size for practical
density values. In case of employing $GF(2)$, the
maximum plausible density is restricted up to 0.5 since higher values incur
in more frequent \ac{l.d.} coded packets \textbf{Include reference here}.

@@

Provided a desired budget $B$ and the \ac{dof} of the decoder, an encoder
can estimate the required coding density for the packets to not exceed the
budget. In our implementation, we use feedback packets to provide the
encoder an estimate of the decoder \ac{dof} at pre-defined points of the
transmission process. The points in the process occur when a decoder
obtains a given amount of \ac{dof}. In our implementation, we define $r(k),\
k \in [1, \ldots)$ as the rank $r$ of the decoder\footnote{The rank of
a decoder is the rank of its coding matrix at a given point in the
transmissions process.} where a decoder has to send the $k$-th feedback in
\eqref{eq:feedback_dofs}.
%
\begin{equation}\label{eq:feedback_dofs}
    r(k) = \floor{g\cdot \left[\frac{2^k -1}{2^k}\right]},\ k \in [1,\ldots)
\end{equation}
%
At the beginning of the transmission process, we assume that the encoder
starts with an initial coding density that has been calculated depending to
the budget. According to \eqref{eq:feedback_dofs}, a decoder sends feedback
once it has obtained (rounded down) $g/2,\ 3g/4,\ 7g/8,\ \ldots$  degrees of
freedom. The total amount of feedback for this scheme will depend on the
generation size. Still, we have an implementation that limits the amount of
feedback to be sent from the decoder in [\textbf{reference}].

The transmissions made between feedback packets are called the density
regions since the same estimated density is used for encoding sparse
packets. Once new feedback is received, an encoder recalculates the coding
density for the new region until the next feedback packet is received.
Before the coding density can be estimated, it is essential that the
encoder has a good estimate of the decoder \ac{dof}, but also the remainder
of the total budget. The feedback scheme based on the rank reports of
\eqref{eq:feedback_dofs} roughly splits the remaining coding regions in
two halves in every. In other words, the first half is the region where
the encoder currently operates. Once an encoder finishes with this region,
it proceeds with second half where again it splits this into to two region
and so on until it finishes on transmitting the whole generation. This is
also the case for the budget that is split equally among the two new
regions. The very last region will be assigned the full remainder of
the total budget and the coding density will not vary.

>>>>

In this section, we present a description of the considered coding schemes
measured with the \ac{Raspi} testbed. We introduce a definition
for the primitive coding operations, e.g. encoding, decoding and recoding
(where applies) for each coding scheme. Later, we address particular schemes
which are obtained by modifying the basic coding operations that provide
better processing speeds, particularly relevant for the \ac{Raspi}.
Finally, we include a review of the \todot{this sentence is incomplete}

@@

\ac{RLNC} is an example of intra-session \ac{NC}. Here, data symbols from a
single flow are combined with each other. In this type of network coding, all
the original data packets $P_j, j \in [1,g]$, and each of $B$ bytes, are used to
create coded packets. In the following subsections, we describe the basic
functionalities of \ac{RLNC}.

\subsubsection{Encoding}
In \ac{RLNC}, any coded packet is a linear combination from all
the original packets. For the coding scheme, packets are seen as
algebraic entities formed as a sequence of elements from $GF(q)$,
which is a \ac{GF} of size $q$. Later, each original packet is
multiplied by a coding coefficient from $GF(q)$. The coding coefficients
are chosen uniformly at random from the \ac{GF} by the encoder. To
perform the multiplication of a packet by a coding coefficient, the
coefficient is multiplied for each of the elements in the
concatenation that composes an original packet, preserving the
concatenation. Later, all resulting packets are added within the
\ac{GF} arithmetics together to generate a coded packet. Thus, a
coded packet can be written as:
%
\begin{align} \label{eq:coded_packet}
C_i  &= \bigoplus_{j=1}^{g} v_{ij} \otimes P_j ,\ \forall i \in [1,\ldots)
\end{align}

@@

For indicating a receiver which packets were utilized
to create a coded one, a simple yet versatile choice is to append
its coding coefficients as a header of the coded packet. Hence, an
amount of overhead is included in every coded packet given that we
need to provide some signaling needed for decoding. The coding
coefficients overhead amount for packet $i$, $|v_i|$, can be quantified
as:
%
\begin{align} \label{eq:coded_packet_coef}
|v_i| &= \sum_{j=1}^{g} |v_{ij}| = g \times \lceil \log_{2} (q)
\rceil ~ [\mathrm{bits}]
\end{align}

\subsubsection{Decoding}
\label{sssec:decoding}
To be able to decode a batch of $g$ packets, a \ac{l.i.} set of $g$
coded packets, $C_i,\ i \in [1,g]$ is required at a decoder.
Once this set has been collected for a decoder,
the original packets can be found by computing the solution of a system
of linear equations using \ac{GF} aritmethics. Thus, we define
$\textbf{C} = \left[C_1 \ldots C_g \right]^T$,
$\textbf{P} = \left[P_1 \ldots P_g \right]^T$ and the coding matrix
$\textbf{V}$ that collects the coding coefficients for each of the $g$
coded packets, as follows:
%
\begin{align} \label{eq:coding_matrix}
\textbf{V} =
\left[
\begin{array}{c}
        v_1    \\ \hline
        \vdots \\ \hline
        v_g    \\
\end{array}
\right]
=
\left[
\begin{array}{ccc}
        v_{11} & \ldots & v_{1g} \\
        \vdots  & \ddots & \vdots  \\
        v_{g1} & \ldots & v_{gg} \\
\end{array}
\right]
\end{align}
%
Algebraically, decoding simplifies to finding the inverse of
$\textbf{V}$ in the linear system
$\textbf{C} = \textbf{V} \textbf{P}$, which can achieved using
efficient Gaussian elimination techniques \cite{fragouli2006network}.
On real applications, decoding is performed on-the-fly, e.g., the pivots are
computed as packets are progressively received, in order to minimize the
computation delay for each step.

@@

In \eqref{eq:recoded_packet}, $w_{ij}$ is the coding coefficient used
in the $i$-th recoded packet and assigned to multiply a previously coded
packet $C_j$. These coding coefficients are again uniformly and randomly chosen
from $GF(q)$. Still, each of the $w_{ij}$ are not the appended recoding
coefficients for signalling since the they will only retrieve previously
coded packets. The analytical expression for the appended coefficients
after recoding is obtained by calculating recoded packets as function of the
original packets. Therefore, we define the local coding matrix $\textbf{W}$
in the same way as it was made for $\textbf{V}$. Thus, the local coding matrix
can be written as:
%
\begin{align} \label{eq:local_coding_matrix}
\textbf{W} =
\left[
\begin{array}{c}
        w_1    \\ \hline
        \vdots \\ \hline
        w_g    \\
\end{array}
\right]
=
\left[
\begin{array}{ccc}
        w_{11} & \ldots & w_{1g} \\
        \vdots  & \ddots & \vdots  \\
        w_{g1} & \ldots & w_{gg} \\
\end{array}
\right]
\end{align}
%
With the definitions from \eqref{eq:coding_matrix} and
\eqref{eq:local_coding_matrix}, the recoded packets
$\textbf{R} = \left[R_1 \ldots R_g \right]^T$ are written as
$\textbf{R} = (\textbf{W} \textbf{V}) \textbf{P}$. Here, we recognize
the relationship between original and recoded packets. The resulting
coding matrix is simply the multiplication of matrices $\textbf{W}$ and
$\textbf{V}$. Denoting ${(\textbf{W} \textbf{V})}_{ij}$ as the element
in the $i$-th row and $j$-th column of $(\textbf{W} \textbf{V})$, this
term is the resulting coding coefficient used to create $R_i$ after
encoding the original packet $P_j$ recoding it locally in an intermediate
node. Finally, the appended coefficients for $R_i$ are
${(\textbf{W} \textbf{V})}_{ik}$ with $k \in [1,g]$. By doing some
algebra, each ${(\textbf{W} \textbf{V})}_{ij}$ term can be verified to be
computed as:
%
\begin{align}
\label{eq:appended_coded_coefficients}
{(\textbf{W} \textbf{V})}_{ij} = \sum_{k=1}^{g} w_{ik} v_{kj}\ ,\ \forall i,j \in [1,g] \times [1,g]
\end{align}
%
This procedure is carried by all the recoders in a given network therefore
allowing any decoder to compute the original data after Gaussian elimination,
regardless of the amount of times where recoding was performed. Similarly
as with the encoding operation, any decoder that collects a set of
$g$ \ac{l.i.} recoded packets with their respective coefficients, will
be able to decode the data as mentioned before in Section \ref{sssec:decoding}.

@@

In \ac{SRLNC}, instead of considering all the packets to create a coded packet
as in \ac{RLNC}, an encoder sets more coding coefficients to zero when
generating a coded packet with the purpose of reducing the overall processing.
Decoding is the same as in \ac{RLNC}, but given that the coding matrices are now
sparse, we will have less operations performed in the decoding process.
Recoding, although theoretically possible, is omitted in practical systems since
it is too complex to keep packets sparse after recoding
[\textbf{reference}]\todot{add ref}. In what follows, we describe the coding
scheme with a basic method to produce sparse coded packets and later we indicate
a second one, giving reference to their implementation performance.

@@

To achieve a desired average coding density in the coefficients, for each
of them, we utilize a set of Bernoulli random variables all with parameter
$d$ as it success probability, i.e. $\mathbb{B}_j \sim Bernoulli(d)\ ,\ \forall
j \in [1,g]$. In this way, we can represent a coded packet in \ac{SRLNC} as:
%
\begin{align} \label{eq:sparse_coded_packet}
    C_i  &= \bigoplus_{j=1}^{g} \mathbb{B}_j v_{ij} \otimes P_j ,\ v_{ij}
    \neq 0\
    \forall i \in [1,\ldots) ,\ \
    d \in
    \begin{cases}
        {(0,0.5]} & ,\ q = 2 \\
        {(0,1]}   & ,\ q > 2
    \end{cases}
\end{align}
%
In \eqref{eq:sparse_coded_packet}, we have the requirement for the coding
coefficient to not be zero, since we want to ensure a coding coefficient
is generated for any random trial where $\mathbb{B}_j = 1$. Therefore, in our
implementation of \ac{SRLNC} we exclude the zero element and then pick
uniformly distributed random elements from $GF(q)-\{ 0 \}$. Also, we have
specified a dependency on the field size for practical
density values. In case of employing $GF(2)$, the
maximum plausible density is restricted up to 0.5 since higher values incur
in more frequent \ac{l.d.} coded packets \textbf{Include reference here}.

@@

The difference between $B$ and $g$ is equal to the losses in the channel and the
losses due to linear dependencies. Therefore, in a lossless channel, the budget
is
\begin{equation}\label{eq:budget}
    B(g,d)=\sum^{g-1}_{i=0}\frac{1}{P(i,g,d)},
\end{equation}

where $P(i,n,d)$ is the probability of receiving an innovative coded packet
after receiving $i$ \ac{l.i.} packets with a coding density $d$. In our implementations, we considered the the upper bound for the innovation probability from \cite{feizi2014tunable}, given as
\begin{equation}
    P(i,g,d)\geq 1-(1-d)^{n-i}
\end{equation}

Provided a desired budget $B$ and the \ac{dof} of the decoder, an encoder
can estimate the required coding density for the packets to not exceed the
budget. In our implementation, we use feedback packets to provide the
encoder an estimate of the decoder \ac{dof} at pre-defined points of the
transmission process. The points in the process occur when a decoder
obtains a given amount of \ac{dof}. In our implementation, we define $r(k),\
k \in [1, \ldots)$ as the rank $r$ of the decoder\footnote{The rank of
a decoder is the rank of its coding matrix at a given point in the
transmissions process.} where a decoder has to send the $k$-th feedback in
\eqref{eq:feedback_dofs}.
%
\begin{equation}\label{eq:feedback_dofs}
    r(k) = \floor{g\cdot \left[\frac{2^k -1}{2^k}\right]},\ k \in [1,\ldots)
\end{equation}
%
At the beginning of the transmission process, we assume that the encoder
starts with an initial coding density that has been calculated depending to
the budget. According to \eqref{eq:feedback_dofs}, a decoder sends feedback
once it has obtained (rounded down) $g/2,\ 3g/4,\ 7g/8,\ \ldots$  degrees of
freedom. The total amount of feedback for this scheme will depend on the
generation size. Still, we have an implementation that limits the amount of
feedback to be sent from the decoder in [\textbf{reference}].\todot{@Chres this reference is missing}

The transmissions made between feedback packets are called the density regions
since the same estimated density is used for encoding sparse packets. Once new
feedback is received, an encoder recalculates the coding density for the new
region until the next feedback packet is received. Before the coding density can
be estimated, it is essential that the encoder has a good estimate of the
decoder \ac{dof}, but also the remainder of the total budget. To calculate the
density, we use bisection to estimate a fixed density for the density region
that satisfies the budget for the $k$-th density region:
\begin{equation}
B(s(k-1),r(k),g,d) = \sum^{r(k)}_{i=r(k-1)} \frac{1}{P(i,g,d)} = \sum^{r(k)}_{i=r(k-1)} \frac{1}{1-(1-d)^{g-i}} = \frac{B}{2^k}, k\in [1,...).
\end{equation}
%
% Where the budget is half of the remaining budget for all the density regions
% except the last one, to which it is assigned the remaining of the budget.
The feedback scheme based on the rank reports of \eqref{eq:feedback_dofs}
roughly splits the remaining coding regions in two halves. In other
words, the first half is the region where the encoder currently operates. Once
an encoder finishes with this region, it proceeds with second half where again
it splits this into to two region and so on until it finishes on transmitting
the whole generation. This is also the case for the budget that is split equally
among the two new regions. The very last region will be assigned the full
remainder of the total budget and the coding density will not vary.
